### 2021-03-30 tables/forecast-types.tex

\documentclass{article}
\usepackage{booktabs}
\usepackage{array}
\usepackage{pdflscape}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{colortbl}
\usepackage{float}
\usepackage{longtable}
\usepackage{multirow}
\usepackage{Sweave}
\begin{document}

\begin{longtable}[t]{>{\raggedright\arraybackslash}p{3.5cm}>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{4.5cm}}
\toprule
\textbf{Forecast type} & \textbf{Target type} & \textbf{Representation of the predictive distribution}\\
\midrule
Point forecast & continuous \newline  integer \newline  binary & one single number for the predicted outcome\\
\cmidrule{1-3}\pagebreak[0]
 & continuous \newline  integer & predictive samples \newline   quantiles \newline   closed analytical form\\
\cmidrule{2-3}\nopagebreak
\multirow[t]{-2}{3.5cm}{\raggedright\arraybackslash Probabilistic forecast} & binary & binary probabilities\\
\bottomrule
\end{longtable}

\end{document}