### 2021-03-30 tables/scores-table-summary.Rnw

\documentclass{article}
\usepackage{booktabs}
\usepackage{array}
\usepackage{pdflscape}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{colortbl}
\usepackage{float}
\usepackage{longtable}
\begin{document}

<<table, echo=FALSE, results=tex>>=
options(prompt = "R> ", continue = "+  ", width = 70, useFancyQuotes = FALSE)
library("MASS")
library(data.table)

crps <- list(
  `Metric` = "CRPS (Continuous) ranked probability score", 
  `Target types` = r"(continuous, integer)", 
  `Forecast formats` = r"(closed-form, samples (approximation))",
  `Properties` = "proper scoring rule, global, stable handling of outliers",
  `References` = ""
)

log_score <- list(
  `Metric` = "Log score", 
  `Target types` = r"(continuous, (integer not in scoringutils))", 
  `Forecast formats` = r"(closed-form, samples (approximation))",
  `Properties` = "proper scoring rule, local, unstable for outliers",
  `References` = ""
)

wis <- list(
  Metric = "WIS (Weighted) interval score", 
  `Target types` = r"(continuous, integer)", 
  `Forecast formats` = r"(quantile or interval predictions)",
  `Properties` = "proper scoring rule, global, stable handling of outliers, converges to crps",
  `References` = ""
)

dss <- list(
  `Metric` = "DSS Dawid-Sebastiani score", 
  `Target types` = r"(continuous, integer)", 
  `Forecast formats` = r"(closed-form, samples (approximation))",
  `Properties` = "proper scoring rule, somewhat global, somewhat stable handling of outliers",
  `References` = ""
)

brier_score <- list(
  `Metric` = "Brier score", 
  `Target types` = r"(binary)", 
  `Forecast formats` = r"(binary probabilities)",
  `Properties` = "proper scoring rule",
  `References` = ""
)

interval_coverage <- list(
  `Metric` = "Interval coverage", 
  `Target types` = r"(continuous, integer)", 
  `Forecast formats` = r"(interval forecasts (needs matching quantiles))",
  `Properties` = "measure for calibration",
  `References` = ""
)

quantile_coverage <- list(
  `Metric` = "Quantile coverage", 
  `Target types` = r"(continuous, integer)", 
  `Forecast formats` = r"(quantile or interval forecasts)",
  `Properties` = "measure for calibration",
  `References` = ""
)

sharpness <- list(
  `Metric` = "Sharpness", 
  `Target types` = r"(continuous, integer)", 
  `Forecast formats` = r"(closed-form, samples, quantile or interval forecasts)",
  `Properties` = "measures sharpness, slightly different depending on forecast format",
  `References` = ""
)

bias <- list(
  `Metric` = "Bias", 
  `Target types` = r"(continuous, integer, quantile)", 
  `Forecast formats` = r"(closed-form, samples, quantile or interval forecasts)",
  `Properties` = "captures tendency to over-or underpredict (aspect of calibration)",
  `References` = ""
)

pit <- list(
  `Metric` = "Probability integral transform (PIT)",
  `Target types` = r"(continuous, integer, quantile)", 
  `Forecast formats` = r"(closed-form, samples, quantile or interval forecasts)",
  `Properties` = "assesses calibration",
  `References` = ""
)

mean_score_ratio <- list(
  `Metric` = "Mean score ratio", 
  `Target types` = r"(depends on score)", 
  `Forecast formats` = r"(depends on score)",
  `Properties` = "compares performance of two models",
  `References` = ""
)

relative_skill <- list(
  `Metric` = "Relative skill", 
  `Target types` = r"(depends on scored)", 
  `Forecast formats` = r"(depends on score)",
  `Properties` = "Ranks models based on pairwise comparisons",
  `References` = ""
)

data <- rbind(as.data.table(crps), 
              as.data.table(log_score), 
              as.data.table(wis), 
              as.data.table(dss), 
              as.data.table(brier_score), 
              as.data.table(interval_coverage),
              as.data.table(quantile_coverage), 
              as.data.table(pit),
              as.data.table(sharpness),
              as.data.table(bias), 
              as.data.table(mean_score_ratio), 
              as.data.table(relative_skill))

saveRDS(data, file = "tables/scores-table-summary.rds")

library(kableExtra)
library(magrittr)
library(knitr)
data[, 1:4] %>%
  kableExtra::kbl(format = "latex", booktabs = TRUE, 
                  escape = FALSE,
                  longtable = TRUE,
                  linesep = c('\\addlinespace')) %>%
  kableExtra::column_spec(1, width = "2.5cm") %>%
  kableExtra::column_spec(2, width = "2cm") %>%
  kableExtra::column_spec(3, width = "4.5cm") %>%
  kableExtra::column_spec(4, width = "4cm") %>%
  kableExtra::kable_styling(latex_options = c("striped", "repeat_header"))

@

\end{document}