### 2021-03-30 tables/scores-table-summary.tex

\documentclass{article}
\usepackage{booktabs}
\usepackage{array}
\usepackage{pdflscape}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{xcolor}
% \usepackage{colortbl}
\usepackage{float}
\usepackage{longtable}
\usepackage{Sweave}
\begin{document}

\begin{longtable}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{4.5cm}>{\raggedright\arraybackslash}p{4cm}}
\toprule
Metric & Target types & Forecast formats & Properties\\
\midrule
\endfirsthead
\multicolumn{4}{@{}l}{\textit{(continued)}}\\
\toprule
Metric & Target types & Forecast formats & Properties\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
\cellcolor{gray!6}{CRPS (Continuous) ranked probability score} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{closed-form, samples (approximation)} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers}\\
\addlinespace
Log score & continuous, (integer not in scoringutils) & closed-form, samples (approximation) & proper scoring rule, local, unstable for outliers\\
\addlinespace
\cellcolor{gray!6}{WIS (Weighted) interval score} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{quantile or interval predictions} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers, converges to crps}\\
\addlinespace
DSS Dawid-Sebastiani score & continuous, integer & closed-form, samples (approximation) & proper scoring rule, somewhat global, somewhat stable handling of outliers\\
\addlinespace
\cellcolor{gray!6}{Brier score} & \cellcolor{gray!6}{binary} & \cellcolor{gray!6}{binary probabilities} & \cellcolor{gray!6}{proper scoring rule}\\
\addlinespace
Interval coverage & continuous, integer & interval forecasts (needs matching quantiles) & measure for calibration\\
\addlinespace
\cellcolor{gray!6}{Quantile coverage} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{quantile or interval forecasts} & \cellcolor{gray!6}{measure for calibration}\\
\addlinespace
Probability integral transform (PIT) & continuous, integer, quantile & closed-form, samples, quantile or interval forecasts & assesses calibration\\
\addlinespace
\cellcolor{gray!6}{Sharpness} & \cellcolor{gray!6}{continuous, integer} & \cellcolor{gray!6}{closed-form, samples, quantile or interval forecasts} & \cellcolor{gray!6}{measures sharpness, slightly different depending on forecast format}\\
\addlinespace
Bias & continuous, integer, quantile & closed-form, samples, quantile or interval forecasts & captures tendency to over-or underpredict (aspect of calibration)\\
\addlinespace
\cellcolor{gray!6}{Mean score ratio} & \cellcolor{gray!6}{depends on score} & \cellcolor{gray!6}{depends on score} & \cellcolor{gray!6}{compares performance of two models}\\
\addlinespace
Relative skill & depends on scored & depends on score & Ranks models based on pairwise comparisons\\*
\end{longtable}

\end{document}