### 2021-11-22 tables/scores-table-summary.tex

\documentclass{article}
\usepackage{booktabs}
\usepackage{array}
\usepackage{pdflscape}
\usepackage{amsfonts}
\usepackage{amsmath}
\usepackage{xcolor}
\usepackage{colortbl}
\usepackage{float}
\usepackage{longtable}
\usepackage{Sweave}
\begin{document}

\begin{longtable}[t]{>{\raggedright\arraybackslash}p{2.5cm}>{\raggedright\arraybackslash}p{2cm}>{\raggedright\arraybackslash}p{4.5cm}>{\raggedright\arraybackslash}p{4cm}}
\toprule
Metric & Target types & Forecast formats & Properties\\
\midrule
\endfirsthead
\multicolumn{4}{@{}l}{\textit{(continued)}}\\
\toprule
Metric & Target types & Forecast formats & Properties\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
\cellcolor{gray!6}{(Continuous) ranked probability score (CRPS)} & \cellcolor{gray!6}{continuous, discrete} & \cellcolor{gray!6}{closed-form, samples (approximation)} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers}\\
\addlinespace
Log score (logS) & continuous, (discrete not in scoringutils) & closed-form, samples (approximation) & proper scoring rule, local, unstable for outliers\\
\addlinespace
\cellcolor{gray!6}{(Weighted) interval score (WIS)} & \cellcolor{gray!6}{continuous, discrete} & \cellcolor{gray!6}{quantile or interval predictions} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers, converges to crps}\\
\addlinespace
Dawid-Sebastiani score (DSS) & continuous, discrete & closed-form, samples (approximation) & proper scoring rule, somewhat global, somewhat stable handling of outliers\\
\addlinespace
\cellcolor{gray!6}{Brier score (BS)} & \cellcolor{gray!6}{binary} & \cellcolor{gray!6}{binary probabilities} & \cellcolor{gray!6}{proper scoring rule}\\
\addlinespace
Interval coverage & continuous, discrete & interval forecasts (needs matching quantiles) & measure for calibration\\
\addlinespace
\cellcolor{gray!6}{Quantile coverage} & \cellcolor{gray!6}{continuous, discrete} & \cellcolor{gray!6}{quantile or interval forecasts} & \cellcolor{gray!6}{measure for calibration}\\
\addlinespace
Probability integral transform (PIT) & continuous, discrete, quantile & closed-form, samples, quantile or interval forecasts & assesses calibration\\
\addlinespace
\cellcolor{gray!6}{Sharpness} & \cellcolor{gray!6}{continuous, discrete} & \cellcolor{gray!6}{closed-form, samples, quantile or interval forecasts} & \cellcolor{gray!6}{measures sharpness, slightly different depending on forecast format}\\
\addlinespace
Bias & continuous, discrete, quantile & closed-form, samples, quantile or interval forecasts & captures tendency to over-or underpredict (aspect of calibration)\\
\addlinespace
\cellcolor{gray!6}{Mean score ratio} & \cellcolor{gray!6}{depends on score} & \cellcolor{gray!6}{depends on score} & \cellcolor{gray!6}{compares performance of two models}\\
\addlinespace
Relative skill & depends on scored & depends on score & Ranks models based on pairwise comparisons\\*
\end{longtable}
\begin{landscape}
\begin{longtable}[t]{>{\raggedright\arraybackslash}p{3.2cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.3cm}>{\centering\arraybackslash}p{1.3cm}>{\centering\arraybackslash}p{1.5cm}>{\centering\arraybackslash}p{1.3cm}>{\raggedright\arraybackslash}p{6.0cm}}
\toprule
Metric & Discrete & Continuous & Binary & Closed-form & Samples (approx.) & Quantiles & Properties\\
\midrule
\endfirsthead
\multicolumn{8}{@{}l}{\textit{(continued)}}\\
\toprule
Metric & Discrete & Continuous & Binary & Closed-form & Samples (approx.) & Quantiles & Properties\\
\midrule
\endhead

\endfoot
\bottomrule
\endlastfoot
\cellcolor{gray!6}{(Continuous) ranked probability score (CRPS)} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers}\\
\addlinespace
Log score (logS) & ($\checkmark$) & $\checkmark$ &  & $\checkmark$ & $\checkmark$ &  & proper scoring rule, local, unstable for outliers\\
\addlinespace
\cellcolor{gray!6}{(Weighted) interval score (WIS)} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{THINK} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{proper scoring rule, global, stable handling of outliers, converges to crps}\\
\addlinespace
Dawid-Sebastiani score (DSS) & $\checkmark$ & $\checkmark$ &  & $\checkmark$ & $\checkmark$ &  & proper scoring rule, somewhat global, somewhat stable handling of outliers\\
\addlinespace
\cellcolor{gray!6}{Brier score (BS)} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{proper scoring rule}\\
\addlinespace
Interval coverage & $\checkmark$ & $\checkmark$ &  &  & THINK & $\checkmark$ & measure for calibration\\
\addlinespace
\cellcolor{gray!6}{Quantile coverage} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{THINK} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{measure for calibration}\\
\addlinespace
Probability integral transform (PIT) & $\checkmark$ & $\checkmark$ &  & $\checkmark$ & $\checkmark$ & $\checkmark$ & assesses calibration\\
\addlinespace
\cellcolor{gray!6}{Sharpness} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{$\checkmark$} & \cellcolor{gray!6}{measures forecast dispersions}\\
\addlinespace
Bias & $\checkmark$ & $\checkmark$ & $\checkmark$??? & $\checkmark$ & $\checkmark$ & $\checkmark$ & captures tendency to over-or underpredict (aspect of calibration)\\
\addlinespace
\cellcolor{gray!6}{Mean score ratio} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{depends} & \cellcolor{gray!6}{compares performance of two models}\\
\addlinespace
Relative skill & depends & depends & depends & depends & depends & depends & Ranks models based on pairwise comparisons\\*
\end{longtable}
\end{landscape}

\end{document}